---
title: "Missingness Imputation"
author: "Johannes Schwenke"
date: today
format:
  typst: default
---

## Overview

This document creates the analysis-specific multiple-imputation objects for the missingness analyses.

```{r}
#| label: setup
library(here)
library(arrow)
library(mice)
library(tidyverse)
library(tinytable)

source(here("R", "io.R"))

followup_limit_days <- 182L

ensure_dir(here("output", "03_imputation"))
```

```{r}
#| label: load-data
patients <- read_parquet(here("output", "derived", "patients_baseline.parquet"))
raw_data <- read_parquet(here("data", "clean", "analysis_data_long_2026-08-11.parquet"))
questionnaires <- read_parquet(here("output", "derived", "questionnaires.parquet")) |>
  filter(
    !is.na(days_since_first_contact),
    days_since_first_contact >= 0,
    days_since_first_contact <= followup_limit_days
  )
items <- read_parquet(here("output", "derived", "questionnaire_items_long.parquet")) |>
  filter(
    !is.na(days_since_first_contact),
    days_since_first_contact >= 0,
    days_since_first_contact <= followup_limit_days
  )
```

```{r}
#| label: build-mi-data
questionnaire_summaries <- questionnaires |>
  summarise(
    n_questionnaires = dplyr::n(),
    mean_prop_missing_nonstructural = mean(prop_missing_nonstructural),
    max_prop_missing_nonstructural = max(prop_missing_nonstructural),
    mean_n_missing_nonstructural = mean(n_items_missing_nonstructural),
    mean_n_terminal_missing = mean(n_terminal_missing),
    any_short_version = any(form_type == "short_version"),
    mean_days_since_first_contact = mean(days_since_first_contact),
    max_days_since_first_contact = max(days_since_first_contact),
    .by = "pat_id"
  )

appointment_summaries <- raw_data |>
  filter(
    !is.na(days_since_fstcnt),
    days_since_fstcnt >= 0,
    days_since_fstcnt <= followup_limit_days
  ) |>
  count(pat_id, name = "n_appointments")

mi_data <- patients |>
  select(-n_questionnaires, -n_appointments, -ecog_baseline) |>
  rename(ecog_baseline = ecog_baseline_analysis) |>
  left_join(questionnaire_summaries, by = "pat_id") |>
  left_join(appointment_summaries, by = "pat_id") |>
  mutate(
    n_questionnaires = tidyr::replace_na(n_questionnaires, 0L),
    n_appointments = tidyr::replace_na(n_appointments, 0L),
    any_short_version = tidyr::replace_na(any_short_version, FALSE),
    across(
      c(
        "mean_prop_missing_nonstructural",
        "max_prop_missing_nonstructural",
        "mean_n_missing_nonstructural",
        "mean_n_terminal_missing",
        "mean_days_since_first_contact",
        "max_days_since_first_contact"
      ),
      ~ tidyr::replace_na(.x, 0)
    ),
    treatment_plan_group = if_else(treatment_plan_group == "Missing", NA_character_, treatment_plan_group),
    diagnosis_group = if_else(diagnosis_group == "Missing", NA_character_, diagnosis_group)
  ) |> 
  select(-diagnosis, -treatment_plan)
```

```{r}
#| label: run-imputation
mi_ready <- mi_data |>
  mutate(
    ecog_baseline = ordered(ecog_baseline),
    sex = factor(sex),
    reminder = factor(reminder),
    treatment_plan_group = factor(treatment_plan_group),
    diagnosis_group = factor(diagnosis_group)
  )

methods <- mice::make.method(mi_ready)
predictor_matrix <- mice::quickpred(mi_ready, mincor = 0.05, minpuc = 0.25)
predictor_matrix[, "pat_id"] <- 0

methods[c(
  "pat_id", "age", "sex", "language", "marital_status", "reminder",
  "n_questionnaires", "n_appointments",
  "mean_prop_missing_nonstructural", "max_prop_missing_nonstructural",
  "mean_n_missing_nonstructural", "mean_n_terminal_missing",
  "any_short_version", "mean_days_since_first_contact",
  "max_days_since_first_contact"
)] <- ""

methods["ecog_baseline"] <- "polr"
methods["diagnosis_group"] <- "polyreg"
methods["treatment_plan_group"] <- "polyreg"

mids_missingness <- mice::mice(
  data = mi_ready,
  m = 10,
  maxit = 20,
  method = methods,
  predictorMatrix = predictor_matrix,
  seed = 20260228,
  printFlag = FALSE
)
```

```{r}
#| label: complete-imputed-datasets
patient_sets <- mice::complete(mids_missingness, action = "all")

questionnaires_imputed <- vector("list", length(patient_sets))
questionnaire_items_imputed <- vector("list", length(patient_sets))

for (i in seq_along(patient_sets)) {
  completed_patients <- patient_sets[[i]]

  questionnaires_imputed[[i]] <- questionnaires |>
    select(
      -tidyselect::any_of(c("ecog_baseline", "diagnosis_group", "treatment_plan_group"))
    ) |>
    left_join(
      completed_patients |>
        select(
          pat_id,
          ecog_baseline,
          diagnosis_group,
          treatment_plan_group
        ),
      by = "pat_id"
    ) |>
    mutate(
      ecog_baseline = ordered(ecog_baseline),
      diagnosis_group = factor(diagnosis_group),
      treatment_plan_group = factor(treatment_plan_group),
      sex = factor(sex),
      reminder = factor(reminder),
      form_type = factor(form_type)
    )

  questionnaire_items_imputed[[i]] <- items |>
    select(
      -tidyselect::any_of(c("ecog_baseline", "diagnosis_group", "treatment_plan_group"))
    ) |>
    left_join(
      completed_patients |>
        select(
          pat_id,
          ecog_baseline,
          diagnosis_group,
          treatment_plan_group
        ),
      by = "pat_id"
    ) |>
    mutate(
      ecog_baseline = ordered(ecog_baseline),
      diagnosis_group = factor(diagnosis_group),
      treatment_plan_group = factor(treatment_plan_group),
      sex = factor(sex),
      reminder = factor(reminder),
      form_type = factor(form_type)
    )
}

completed_sets <- list(
  patients = patient_sets,
  questionnaires = questionnaires_imputed,
  questionnaire_items = questionnaire_items_imputed
)

saveRDS(mids_missingness, here("output", "03_imputation", "mids_missingness.rds"))
saveRDS(completed_sets$questionnaires, here("output", "03_imputation", "questionnaires_imputed_list.rds"))
saveRDS(completed_sets$questionnaire_items, here("output", "03_imputation", "questionnaire_items_imputed_list.rds"))

# brms can read the mids object directly via brm_multiple(); no additional
# conversion is needed — saving mids_missingness.rds above is sufficient.
```

### Convergence

Trace plots of the per-chain mean and SD over iterations. Chains should mix well
and show no systematic trend.

```{r}
#| label: diag-trace
#| fig-width: 8
#| fig-height: 5
plot(mids_missingness)
```

### Density plots

Observed (blue) vs. imputed (red) marginal distributions for the continuous /
ordinal target variables.

```{r}
#| label: diag-density-ecog
#| fig-width: 6
#| fig-height: 4
densityplot(mids_missingness, ~ ecog_baseline,
  xlab = "ECOG baseline (as integer)"
)
```

### Strip plots

Strip plots show every observed (blue) and imputed (red) value for the
categorical variables, jittered across the  imputed datasets.

```{r}
#| label: diag-strip-diagnosis
#| fig-width: 7
#| fig-height: 4
stripplot(mids_missingness, diagnosis_group ~ .imp,
  xlab = "Imputation index",
  ylab = "Diagnosis group"
)
```

```{r}
#| label: diag-strip-treatment
#| fig-width: 7
#| fig-height: 4
stripplot(mids_missingness, treatment_plan_group ~ .imp,
  xlab = "Imputation index",
  ylab = "Treatment plan group"
)
```
