---
title: "EORTC Score Availability"
author: "Johannes Schwenke"
date: today
format:
  typst: default
bibliography: ../references.bib
---

## Overview

This document calculates how many EORTC scores can be calculated from each recorded questionnaire during the first six months after first contact.

```{r}
#| label: setup
library(here)
library(arrow)
library(tidyverse)
library(scales)
library(tinytable)

source(here("R", "io.R"))

output_dir <- here("output", "04_score_computability")
followup_limit_days <- 182L

ensure_dir(output_dir)
```

## Analysis Data

Questionnaires recorded from day 0 through day 182 are included. Questionnaires with all items missing remain in the analysis and yield no calculated scores.

```{r}
#| label: load-data
questionnaires <- read_parquet(
  here("output", "derived", "questionnaires.parquet")
) |>
  filter(
    !is.na(days_since_first_contact),
    days_since_first_contact >= 0,
    days_since_first_contact <= followup_limit_days
  ) |>
  mutate(
    assessment = if_else(
      days_since_first_contact == 0,
      "First contact",
      "Follow-up"
    ),
    assessment = factor(
      assessment,
      levels = c("First contact", "Follow-up")
    ),
    scoring_rule = if_else(
      assessment == "First contact",
      "QLQ-C30",
      "QLQ-C15-PAL"
    ),
    administered_form = case_when(
      form_type == "long_version" ~ "EORTC QLQ-C30",
      form_type == "short_version" ~
        "Adapted EORTC QLQ-C15-based follow-up form"
    )
  )

questionnaire_items <- read_parquet(
  here("output", "derived", "questionnaire_items_long.parquet")
) |>
  filter(questionnaire_key %in% questionnaires$questionnaire_key) |>
  filter(str_detect(item_code, "^q([1-9]|[12][0-9]|30)$")) |>
  mutate(
    response = case_when(
      item_code %in% c("q29", "q30") & response %in% 1:7 ~ response,
      !(item_code %in% c("q29", "q30")) & response %in% 1:4 ~ response,
      TRUE ~ NA_real_
    )
  ) |>
  select(questionnaire_key, item_code, response)

questionnaire_items_wide <- questionnaire_items |>
  pivot_wider(
    names_from = item_code,
    values_from = response
  )

score_data <- questionnaires |>
  left_join(questionnaire_items_wide, by = "questionnaire_key")
```

## Item and Score Crosswalk

Stored item names always use QLQ-C30 numbering. The crosswalk records the corresponding QLQ-C15-PAL item numbers used for follow-up scoring.

```{r}
#| label: score-crosswalk
c30_score_item_crosswalk <- tribble(
  ~scoring_rule, ~score_code, ~score_label, ~score_type, ~c15_item_number, ~c30_item_code, ~n_items_total, ~n_items_required,
  "QLQ-C30", "QL2", "Global health status / quality of life", "global", NA_integer_, "q29", 2L, 1L,
  "QLQ-C30", "QL2", "Global health status / quality of life", "global", NA_integer_, "q30", 2L, 1L,
  "QLQ-C30", "PF2", "Physical functioning", "function", NA_integer_, "q1", 5L, 3L,
  "QLQ-C30", "PF2", "Physical functioning", "function", NA_integer_, "q2", 5L, 3L,
  "QLQ-C30", "PF2", "Physical functioning", "function", NA_integer_, "q3", 5L, 3L,
  "QLQ-C30", "PF2", "Physical functioning", "function", NA_integer_, "q4", 5L, 3L,
  "QLQ-C30", "PF2", "Physical functioning", "function", NA_integer_, "q5", 5L, 3L,
  "QLQ-C30", "RF2", "Role functioning", "function", NA_integer_, "q6", 2L, 1L,
  "QLQ-C30", "RF2", "Role functioning", "function", NA_integer_, "q7", 2L, 1L,
  "QLQ-C30", "EF", "Emotional functioning", "function", NA_integer_, "q21", 4L, 2L,
  "QLQ-C30", "EF", "Emotional functioning", "function", NA_integer_, "q22", 4L, 2L,
  "QLQ-C30", "EF", "Emotional functioning", "function", NA_integer_, "q23", 4L, 2L,
  "QLQ-C30", "EF", "Emotional functioning", "function", NA_integer_, "q24", 4L, 2L,
  "QLQ-C30", "CF", "Cognitive functioning", "function", NA_integer_, "q20", 2L, 1L,
  "QLQ-C30", "CF", "Cognitive functioning", "function", NA_integer_, "q25", 2L, 1L,
  "QLQ-C30", "SF", "Social functioning", "function", NA_integer_, "q26", 2L, 1L,
  "QLQ-C30", "SF", "Social functioning", "function", NA_integer_, "q27", 2L, 1L,
  "QLQ-C30", "FA", "Fatigue", "symptom", NA_integer_, "q10", 3L, 2L,
  "QLQ-C30", "FA", "Fatigue", "symptom", NA_integer_, "q12", 3L, 2L,
  "QLQ-C30", "FA", "Fatigue", "symptom", NA_integer_, "q18", 3L, 2L,
  "QLQ-C30", "NV", "Nausea and vomiting", "symptom", NA_integer_, "q14", 2L, 1L,
  "QLQ-C30", "NV", "Nausea and vomiting", "symptom", NA_integer_, "q15", 2L, 1L,
  "QLQ-C30", "PA", "Pain", "symptom", NA_integer_, "q9", 2L, 1L,
  "QLQ-C30", "PA", "Pain", "symptom", NA_integer_, "q19", 2L, 1L,
  "QLQ-C30", "DY", "Dyspnoea", "symptom", NA_integer_, "q8", 1L, 1L,
  "QLQ-C30", "SL", "Insomnia", "symptom", NA_integer_, "q11", 1L, 1L,
  "QLQ-C30", "AP", "Appetite loss", "symptom", NA_integer_, "q13", 1L, 1L,
  "QLQ-C30", "CO", "Constipation", "symptom", NA_integer_, "q16", 1L, 1L,
  "QLQ-C30", "DI", "Diarrhoea", "symptom", NA_integer_, "q17", 1L, 1L,
  "QLQ-C30", "FI", "Financial difficulties", "symptom", NA_integer_, "q28", 1L, 1L
)

c15_score_item_crosswalk <- tribble(
  ~scoring_rule, ~score_code, ~score_label, ~score_type, ~c15_item_number, ~c30_item_code, ~n_items_total, ~n_items_required,
  "QLQ-C15-PAL", "PF2", "Physical functioning", "function", 1L, "q3", 3L, 3L,
  "QLQ-C15-PAL", "PF2", "Physical functioning", "function", 2L, "q4", 3L, 3L,
  "QLQ-C15-PAL", "PF2", "Physical functioning", "function", 3L, "q5", 3L, 3L,
  "QLQ-C15-PAL", "DY", "Dyspnoea", "symptom", 4L, "q8", 1L, 1L,
  "QLQ-C15-PAL", "PA", "Pain", "symptom", 5L, "q9", 2L, 1L,
  "QLQ-C15-PAL", "SL", "Insomnia", "symptom", 6L, "q11", 1L, 1L,
  "QLQ-C15-PAL", "FA", "Fatigue", "symptom", 7L, "q12", 2L, 2L,
  "QLQ-C15-PAL", "AP", "Appetite loss", "symptom", 8L, "q13", 1L, 1L,
  "QLQ-C15-PAL", "NV", "Nausea and vomiting", "symptom", 9L, "q14", 1L, 1L,
  "QLQ-C15-PAL", "CO", "Constipation", "symptom", 10L, "q16", 1L, 1L,
  "QLQ-C15-PAL", "FA", "Fatigue", "symptom", 11L, "q18", 2L, 2L,
  "QLQ-C15-PAL", "PA", "Pain", "symptom", 12L, "q19", 2L, 1L,
  "QLQ-C15-PAL", "EF", "Emotional functioning", "function", 13L, "q21", 2L, 2L,
  "QLQ-C15-PAL", "EF", "Emotional functioning", "function", 14L, "q24", 2L, 2L,
  "QLQ-C15-PAL", "QL", "Quality of life", "global", 15L, "q30", 1L, 1L
)

score_item_crosswalk <- bind_rows(
  c30_score_item_crosswalk,
  c15_score_item_crosswalk
)

write_parquet(
  score_item_crosswalk,
  here(output_dir, "score_item_crosswalk.parquet")
)

score_rule_table <- score_item_crosswalk |>
  arrange(scoring_rule, score_code, c15_item_number, c30_item_code) |>
  summarise(
    `Stored items` = paste(c30_item_code, collapse = ", "),
    `C15-PAL items` = if_else(
      all(is.na(c15_item_number)),
      "-",
      paste(c15_item_number, collapse = ", ")
    ),
    Required = first(n_items_required),
    `Total items` = first(n_items_total),
    .by = c(scoring_rule, score_code, score_label)
  ) |>
  rename(
    `Scoring rules` = scoring_rule,
    Score = score_code,
    Measure = score_label
  )

tt(
  score_rule_table,
  caption = "EORTC score and item crosswalk. Stored item names use QLQ-C30 numbering."
)
```

## Score Calculation

### QLQ-C30 rules at first contact

Multi-item scores are calculated when at least half of their items have responses. Missing components are omitted from the raw-score mean [@Fayers2001-eortc-scoring].

```{r}
#| label: calculate-c30-scores
c30_scores <- score_data |>
  filter(assessment == "First contact") |>
  mutate(
    c30_ql2_n = rowSums(!is.na(pick(q29, q30))),
    c30_ql2_raw = rowMeans(pick(q29, q30), na.rm = TRUE),
    c30_ql2 = if_else(
      c30_ql2_n >= 1,
      (c30_ql2_raw - 1) / 6 * 100,
      NA_real_
    ),
    c30_pf2_n = rowSums(!is.na(pick(q1, q2, q3, q4, q5))),
    c30_pf2_raw = rowMeans(pick(q1, q2, q3, q4, q5), na.rm = TRUE),
    c30_pf2 = if_else(
      c30_pf2_n >= 3,
      (1 - (c30_pf2_raw - 1) / 3) * 100,
      NA_real_
    ),
    c30_rf2_n = rowSums(!is.na(pick(q6, q7))),
    c30_rf2_raw = rowMeans(pick(q6, q7), na.rm = TRUE),
    c30_rf2 = if_else(
      c30_rf2_n >= 1,
      (1 - (c30_rf2_raw - 1) / 3) * 100,
      NA_real_
    ),
    c30_ef_n = rowSums(!is.na(pick(q21, q22, q23, q24))),
    c30_ef_raw = rowMeans(pick(q21, q22, q23, q24), na.rm = TRUE),
    c30_ef = if_else(
      c30_ef_n >= 2,
      (1 - (c30_ef_raw - 1) / 3) * 100,
      NA_real_
    ),
    c30_cf_n = rowSums(!is.na(pick(q20, q25))),
    c30_cf_raw = rowMeans(pick(q20, q25), na.rm = TRUE),
    c30_cf = if_else(
      c30_cf_n >= 1,
      (1 - (c30_cf_raw - 1) / 3) * 100,
      NA_real_
    ),
    c30_sf_n = rowSums(!is.na(pick(q26, q27))),
    c30_sf_raw = rowMeans(pick(q26, q27), na.rm = TRUE),
    c30_sf = if_else(
      c30_sf_n >= 1,
      (1 - (c30_sf_raw - 1) / 3) * 100,
      NA_real_
    ),
    c30_fa_n = rowSums(!is.na(pick(q10, q12, q18))),
    c30_fa_raw = rowMeans(pick(q10, q12, q18), na.rm = TRUE),
    c30_fa = if_else(
      c30_fa_n >= 2,
      (c30_fa_raw - 1) / 3 * 100,
      NA_real_
    ),
    c30_nv_n = rowSums(!is.na(pick(q14, q15))),
    c30_nv_raw = rowMeans(pick(q14, q15), na.rm = TRUE),
    c30_nv = if_else(
      c30_nv_n >= 1,
      (c30_nv_raw - 1) / 3 * 100,
      NA_real_
    ),
    c30_pa_n = rowSums(!is.na(pick(q9, q19))),
    c30_pa_raw = rowMeans(pick(q9, q19), na.rm = TRUE),
    c30_pa = if_else(
      c30_pa_n >= 1,
      (c30_pa_raw - 1) / 3 * 100,
      NA_real_
    ),
    c30_dy = if_else(!is.na(q8), (q8 - 1) / 3 * 100, NA_real_),
    c30_sl = if_else(!is.na(q11), (q11 - 1) / 3 * 100, NA_real_),
    c30_ap = if_else(!is.na(q13), (q13 - 1) / 3 * 100, NA_real_),
    c30_co = if_else(!is.na(q16), (q16 - 1) / 3 * 100, NA_real_),
    c30_di = if_else(!is.na(q17), (q17 - 1) / 3 * 100, NA_real_),
    c30_fi = if_else(!is.na(q28), (q28 - 1) / 3 * 100, NA_real_)
  ) |>
  select(
    pat_id,
    questionnaire_number,
    questionnaire_key,
    assessment,
    scoring_rule,
    administered_form,
    form_type,
    noquestionnaire,
    days_since_first_contact,
    app_day,
    encounter_type,
    c30_ql2,
    c30_pf2,
    c30_rf2,
    c30_ef,
    c30_cf,
    c30_sf,
    c30_fa,
    c30_nv,
    c30_pa,
    c30_dy,
    c30_sl,
    c30_ap,
    c30_co,
    c30_di,
    c30_fi
  )
```

### QLQ-C15-PAL rules during follow-up

The shortened PF2, EF, FA, and NV scores use the published QLQ-C15-PAL transformations and require every component item. Pain retains the QLQ-C30 missing-item rule, while each single-item score requires its item [@Groenvold2006-c15-scoring].

```{r}
#| label: calculate-c15-scores
c15_scores <- score_data |>
  filter(assessment == "Follow-up") |>
  mutate(
    c15_pf2_sum = 12 - (q3 + q4 + q5),
    c15_pf2 = case_when(
      is.na(c15_pf2_sum) ~ NA_real_,
      c15_pf2_sum >= 0 & c15_pf2_sum <= 5 ~ c15_pf2_sum * 100 / 15,
      c15_pf2_sum == 6 ~ 7 * 100 / 15,
      c15_pf2_sum == 7 ~ 9 * 100 / 15,
      c15_pf2_sum == 8 ~ 11 * 100 / 15,
      c15_pf2_sum == 9 ~ 14 * 100 / 15,
      TRUE ~ NA_real_
    ),
    c15_fa_sum = q12 + q18 - 2,
    c15_fa = case_when(
      is.na(c15_fa_sum) ~ NA_real_,
      c15_fa_sum == 0 ~ 0,
      c15_fa_sum == 1 ~ 2 * 100 / 9,
      c15_fa_sum == 2 ~ 3 * 100 / 9,
      (q12 == 3 & q18 == 2) | (q12 == 4 & q18 == 1) ~ 4 * 100 / 9,
      (q12 == 1 & q18 == 4) | (q12 == 2 & q18 == 3) ~ 5 * 100 / 9,
      c15_fa_sum == 4 ~ 6 * 100 / 9,
      c15_fa_sum == 5 ~ 8 * 100 / 9,
      c15_fa_sum == 6 ~ 100,
      TRUE ~ NA_real_
    ),
    c15_nv = case_when(
      q14 == 1 ~ 0,
      q14 == 2 ~ 100 / 6,
      q14 == 3 ~ 50,
      q14 == 4 ~ 100,
      TRUE ~ NA_real_
    ),
    c15_ef_sum = 8 - (q21 + q24),
    c15_ef = case_when(
      is.na(c15_ef_sum) ~ NA_real_,
      c15_ef_sum == 0 ~ 0,
      c15_ef_sum == 1 ~ 2 * 100 / 12,
      c15_ef_sum == 2 ~ 5 * 100 / 12,
      c15_ef_sum == 3 ~ 50,
      c15_ef_sum == 4 ~ 8 * 100 / 12,
      c15_ef_sum == 5 ~ 10 * 100 / 12,
      c15_ef_sum == 6 ~ 100,
      TRUE ~ NA_real_
    ),
    c15_pa_n = rowSums(!is.na(pick(q9, q19))),
    c15_pa_raw = rowMeans(pick(q9, q19), na.rm = TRUE),
    c15_pa = if_else(
      c15_pa_n >= 1,
      (c15_pa_raw - 1) / 3 * 100,
      NA_real_
    ),
    c15_dy = if_else(!is.na(q8), (q8 - 1) / 3 * 100, NA_real_),
    c15_sl = if_else(!is.na(q11), (q11 - 1) / 3 * 100, NA_real_),
    c15_ap = if_else(!is.na(q13), (q13 - 1) / 3 * 100, NA_real_),
    c15_co = if_else(!is.na(q16), (q16 - 1) / 3 * 100, NA_real_),
    c15_ql = if_else(!is.na(q30), (q30 - 1) / 6 * 100, NA_real_)
  ) |>
  select(
    pat_id,
    questionnaire_number,
    questionnaire_key,
    assessment,
    scoring_rule,
    administered_form,
    form_type,
    noquestionnaire,
    days_since_first_contact,
    app_day,
    encounter_type,
    c15_pf2,
    c15_ef,
    c15_fa,
    c15_nv,
    c15_pa,
    c15_dy,
    c15_sl,
    c15_ap,
    c15_co,
    c15_ql
  )
```

## Score Availability

```{r}
#| label: derive-score-availability
score_values_wide <- bind_rows(c30_scores, c15_scores)

score_column_lookup <- tribble(
  ~scoring_rule, ~score_column, ~score_code,
  "QLQ-C30", "c30_ql2", "QL2",
  "QLQ-C30", "c30_pf2", "PF2",
  "QLQ-C30", "c30_rf2", "RF2",
  "QLQ-C30", "c30_ef", "EF",
  "QLQ-C30", "c30_cf", "CF",
  "QLQ-C30", "c30_sf", "SF",
  "QLQ-C30", "c30_fa", "FA",
  "QLQ-C30", "c30_nv", "NV",
  "QLQ-C30", "c30_pa", "PA",
  "QLQ-C30", "c30_dy", "DY",
  "QLQ-C30", "c30_sl", "SL",
  "QLQ-C30", "c30_ap", "AP",
  "QLQ-C30", "c30_co", "CO",
  "QLQ-C30", "c30_di", "DI",
  "QLQ-C30", "c30_fi", "FI",
  "QLQ-C15-PAL", "c15_pf2", "PF2",
  "QLQ-C15-PAL", "c15_ef", "EF",
  "QLQ-C15-PAL", "c15_fa", "FA",
  "QLQ-C15-PAL", "c15_nv", "NV",
  "QLQ-C15-PAL", "c15_pa", "PA",
  "QLQ-C15-PAL", "c15_dy", "DY",
  "QLQ-C15-PAL", "c15_sl", "SL",
  "QLQ-C15-PAL", "c15_ap", "AP",
  "QLQ-C15-PAL", "c15_co", "CO",
  "QLQ-C15-PAL", "c15_ql", "QL"
)

score_metadata <- score_item_crosswalk |>
  distinct(
    scoring_rule,
    score_code,
    score_label,
    score_type,
    n_items_total,
    n_items_required
  )

score_answer_counts <- questionnaires |>
  select(questionnaire_key, scoring_rule) |>
  inner_join(
    score_item_crosswalk,
    by = "scoring_rule",
    relationship = "many-to-many"
  ) |>
  left_join(
    questionnaire_items,
    by = c(
      "questionnaire_key",
      "c30_item_code" = "item_code"
    )
  ) |>
  summarise(
    n_items_answered = sum(!is.na(response)),
    .by = c(
      questionnaire_key,
      scoring_rule,
      score_code
    )
  )

score_values_long <- score_values_wide |>
  pivot_longer(
    cols = matches("^(c30|c15)_"),
    names_to = "score_column",
    values_to = "score_value"
  ) |>
  inner_join(
    score_column_lookup,
    by = c("scoring_rule", "score_column")
  ) |>
  left_join(
    score_metadata,
    by = c("scoring_rule", "score_code")
  ) |>
  left_join(
    score_answer_counts,
    by = c("questionnaire_key", "scoring_rule", "score_code")
  ) |>
  mutate(computable = !is.na(score_value)) |>
  select(
    pat_id,
    questionnaire_number,
    questionnaire_key,
    assessment,
    scoring_rule,
    administered_form,
    form_type,
    noquestionnaire,
    days_since_first_contact,
    app_day,
    encounter_type,
    score_code,
    score_label,
    score_type,
    n_items_total,
    n_items_required,
    n_items_answered,
    computable,
    score_value
  ) |>
  arrange(pat_id, questionnaire_number, score_code)

questionnaire_score_availability <- score_values_long |>
  summarise(
    n_scores_possible = n(),
    n_scores_calculated = sum(computable),
    prop_scores_calculated = mean(computable),
    all_scores_calculated = all(computable),
    some_scores_calculated = any(computable) & !all(computable),
    no_scores_calculated = !any(computable),
    .by = c(
      pat_id,
      questionnaire_number,
      questionnaire_key,
      assessment,
      scoring_rule,
      administered_form,
      form_type,
      noquestionnaire,
      days_since_first_contact,
      app_day,
      encounter_type
    )
  ) |>
  arrange(pat_id, questionnaire_number)

score_availability_summary <- questionnaire_score_availability |>
  summarise(
    n_questionnaires = n(),
    n_scores_possible = first(n_scores_possible),
    mean_scores_calculated = mean(n_scores_calculated),
    sd_scores_calculated = sd(n_scores_calculated),
    mean_prop_scores_calculated = mean(prop_scores_calculated),
    n_all_scores_calculated = sum(all_scores_calculated),
    prop_all_scores_calculated = mean(all_scores_calculated),
    n_no_scores_calculated = sum(no_scores_calculated),
    prop_no_scores_calculated = mean(no_scores_calculated),
    .by = assessment
  ) |>
  arrange(assessment)

score_availability_by_measure <- score_values_long |>
  summarise(
    n_questionnaires = n(),
    n_scores_calculated = sum(computable),
    prop_scores_calculated = mean(computable),
    .by = c(
      assessment,
      scoring_rule,
      score_code,
      score_label
    )
  ) |>
  arrange(assessment, score_code)

form_use_summary <- questionnaires |>
  count(assessment, administered_form, name = "n_questionnaires") |>
  arrange(assessment, administered_form)
```

## Validation

```{r}
#| label: validate-score-calculation
stopifnot(
  nrow(c30_score_item_crosswalk) == 30L,
  n_distinct(c30_score_item_crosswalk$score_code) == 15L,
  nrow(c15_score_item_crosswalk) == 15L,
  n_distinct(c15_score_item_crosswalk$score_code) == 10L,
  all(
    questionnaire_score_availability$n_scores_possible[
      questionnaire_score_availability$assessment == "First contact"
    ] == 15L
  ),
  all(
    questionnaire_score_availability$n_scores_possible[
      questionnaire_score_availability$assessment == "Follow-up"
    ] == 10L
  ),
  nrow(questionnaire_score_availability) == nrow(questionnaires),
  all(
    is.na(score_values_long$score_value) |
      (score_values_long$score_value >= 0 & score_values_long$score_value <= 100)
  ),
  all(
    score_values_long$computable ==
      (score_values_long$n_items_answered >= score_values_long$n_items_required)
  )
)

all_missing_questionnaire_keys <- questionnaires |>
  filter(n_items_answered == 0) |>
  pull(questionnaire_key)

stopifnot(
  all(
    questionnaire_score_availability$n_scores_calculated[
      questionnaire_score_availability$questionnaire_key %in%
        all_missing_questionnaire_keys
    ] == 0L
  )
)

c15_shortened_scores <- score_values_long |>
  filter(
    scoring_rule == "QLQ-C15-PAL",
    score_code %in% c("PF2", "EF", "FA", "NV")
  )

stopifnot(
  all(
    c15_shortened_scores$computable ==
      (c15_shortened_scores$n_items_answered ==
        c15_shortened_scores$n_items_total)
  )
)
```

## Saved Outputs

```{r}
#| label: save-score-outputs
write_parquet(
  score_values_long,
  here(output_dir, "score_values_long.parquet")
)
write_parquet(
  score_values_wide,
  here(output_dir, "score_values_wide.parquet")
)
write_parquet(
  questionnaire_score_availability,
  here(output_dir, "questionnaire_score_availability.parquet")
)
write_parquet(
  score_availability_summary,
  here(output_dir, "score_availability_summary.parquet")
)
write_parquet(
  score_availability_by_measure,
  here(output_dir, "score_availability_by_measure.parquet")
)
```

## Summary Results

```{r}
#| label: tbl-score-availability-summary
#| tbl-cap: "Availability of EORTC scores at first contact and during follow-up."
score_availability_display <- score_availability_summary |>
  mutate(
    Assessment = as.character(assessment),
    Questionnaires = n_questionnaires,
    `Scores calculated, mean/max (%)` = paste0(
      number(mean_scores_calculated, accuracy = 0.1),
      "/",
      n_scores_possible,
      " (",
      percent(mean_prop_scores_calculated, accuracy = 0.1),
      ")"
    ),
    `All scores calculated, n (%)` = paste0(
      n_all_scores_calculated,
      " (",
      percent(prop_all_scores_calculated, accuracy = 0.1),
      ")"
    ),
    `No scores calculated, n (%)` = paste0(
      n_no_scores_calculated,
      " (",
      percent(prop_no_scores_calculated, accuracy = 0.1),
      ")"
    )
  ) |>
  select(
    Assessment,
    Questionnaires,
    `Scores calculated, mean/max (%)`,
    `All scores calculated, n (%)`,
    `No scores calculated, n (%)`
  )

tt(score_availability_display)
```

```{r}
#| label: tbl-score-availability-measure
#| tbl-cap: "Availability of individual EORTC scores at first contact and during follow-up."
score_availability_measure_display <- score_availability_by_measure |>
  mutate(
    Assessment = as.character(assessment),
    Score = score_code,
    Measure = score_label,
    `Calculated, n (%)` = paste0(
      n_scores_calculated,
      " (",
      percent(prop_scores_calculated, accuracy = 0.1),
      ")"
    )
  ) |>
  select(Assessment, Score, Measure, `Calculated, n (%)`)

tt(score_availability_measure_display)
```

## Form Use Check

This internal check shows which questionnaire form was administered at each assessment. It does not change the assessment-specific scoring rules.

```{r}
#| label: tbl-form-use-check
tt(form_use_summary)
```
