---
title: "Survey Completion Time"
author: "Johannes Schwenke"
date: today
format:
  typst: default
---

## Overview

This document summarizes the time required to complete or partially complete the long and short questionnaires. Records without a positive survey time are not included. Questionnaire form is classified from encounter type using the same rule as the shared analysis-dataset pipeline.

```{r}
#| label: setup
library(here)
library(arrow)
library(tidyverse)
library(tinytable)

source(here("R", "io.R"))

output_dir <- here("output", "07_survey_completion_time")
ensure_dir(output_dir)

followup_limit_days <- 182L
```

## Analysis Data

Survey time is recorded in minutes. Questionnaires recorded from day 0 through day 182 after first contact are included. All positive recorded durations within this follow-up window are included without truncation, exclusion, or imputation. Median and interquartile range are used because the duration distribution is strongly right-skewed and includes surveys that may have remained open after active completion stopped.

```{r}
#| label: load-data
survey_time_source <- read_parquet(
  here("data", "clean", "analysis_data_long_2026-08-11.parquet")
) |>
  filter(
    quest_status == 1,
    encounter_type %in% c(1, 2, 3),
    !is.na(days_since_fstcnt),
    days_since_fstcnt >= 0,
    days_since_fstcnt <= followup_limit_days
  ) |>
  mutate(
    form_type = case_when(
      encounter_type == 1 ~ "long_version",
      encounter_type %in% c(2, 3) ~ "short_version"
    ),
    questionnaire_form = case_when(
      form_type == "long_version" ~ "Long questionnaire",
      form_type == "short_version" ~ "Short questionnaire"
    ),
    questionnaire_form = factor(
      questionnaire_form,
      levels = c("Long questionnaire", "Short questionnaire")
    )
  )

survey_time_data <- survey_time_source |>
  filter(!is.na(survey_time), survey_time > 0)
```

## Completion-Time Summary

```{r}
#| label: summarize-survey-time
survey_time_summary <- survey_time_data |>
  summarise(
    n = n(),
    median_minutes = median(survey_time),
    q1_minutes = quantile(survey_time, 0.25),
    q3_minutes = quantile(survey_time, 0.75),
    .by = questionnaire_form
  ) |>
  arrange(questionnaire_form)

survey_time_data_quality <- survey_time_source |>
  summarise(
    n_eligible_questionnaire_records = n(),
    n_positive_recorded_durations = sum(
      !is.na(survey_time) & survey_time > 0
    ),
    n_without_positive_recorded_duration = sum(
      is.na(survey_time) | survey_time <= 0
    ),
    .by = questionnaire_form
  ) |>
  arrange(questionnaire_form)

survey_time_metadata <- tibble(
  name = c(
    "n_positive_recorded_durations",
    "summary_measure"
  ),
  value = c(
    as.character(nrow(survey_time_data)),
    "Median and interquartile range"
  )
)

write_parquet(
  survey_time_summary,
  here(output_dir, "survey_time_summary.parquet")
)
write_parquet(
  survey_time_data_quality,
  here(output_dir, "survey_time_data_quality.parquet")
)
write_parquet(
  survey_time_metadata,
  here(output_dir, "metadata.parquet")
)
```

```{r}
#| label: tbl-survey-time-summary
#| tbl-cap: "Survey completion time by questionnaire form."

survey_time_table <- survey_time_summary |>
  mutate(
    `Median [Q1, Q3], minutes` = sprintf(
      "%.1f [%.1f, %.1f]",
      median_minutes,
      q1_minutes,
      q3_minutes
    )
  ) |>
  select(
    `Questionnaire form` = questionnaire_form,
    n,
    `Median [Q1, Q3], minutes`
  )

tt(survey_time_table)
```

## Data Availability

```{r}
#| label: tbl-survey-time-data-quality
#| tbl-cap: "Availability of positive recorded survey durations."

survey_time_data_quality_table <- survey_time_data_quality |>
  select(
    `Questionnaire form` = questionnaire_form,
    `Eligible questionnaire records` = n_eligible_questionnaire_records,
    `Positive recorded durations` = n_positive_recorded_durations,
    `Without positive recorded duration` = n_without_positive_recorded_duration
  )

tt(survey_time_data_quality_table)
```
